prolfqua&colon; A Comprehensive R-package for Proteomics Differential Expression Analysis

### 2022-10-20 inst/MyArticle/paper.Rmd
Edits in method section.

The linear model below explain the observed protein abundances using the explanatory variables `condition_` and `batch_` and the interaction among them `condition_:batch_`,

```{r, echo = TRUE}
data_Yeast2Factor <- prolfqua::prolfqua_data("data_Yeast2Factor")
mlm <- prolfqua::strategy_lm(transformedIntensity ~ condition_ + batch_ + condition_:batch_)
mm <- prolfqua::build_model(data_Yeast2Factor$data, mlm, subject = "protein_Id")
```

then the contrasts among the groups `Ethanol` and `Glucose` defined by the factor `condition`, and similarly for the difference of the groups defined by `batch_`, can be specified as:

@@

Furthermore, the difference among `Ethanol` and `Glucose` within the batches `p2370` and `p2691` can be specified as follows:

@@

Finally, we can examine if the differences among `conditions_` observed in both batches significantly differ by specifying the contrast:

@@

The code to compute the array of weights $c$ used to multiply the model coefficient $\beta$, to obtain the effect size  estimates, is shown below:

@@

while Table \@ref(tab:contrcoefficient) shows the weight $c$ of each model parameter in each contrast.

@@

This example is taken from the `r BiocStyle::Githubpkg("fgcz/prolfqua")` package vignette file.
Given the coefficients, $c$ contrasts for models obtained with `r BiocStyle::Biocpkg('proDA')`, or `r BiocStyle::Biocpkg('msqrob2')` can be determined, enabling us to integrate these packages transparently.

@@

## compute contrasts from linear model
contr <- prolfqua::Contrasts$new(models, contrasts) 
# compute contrasts using imputation imputation
conI <- prolfqua::ContrastsSimpleImpute$new( lfqp, contrasts)
# merge contrasts, to obtain differences for all proteins
contrasts <- prolfqua::merge_contrasts_results(prefer = contr, add = conI)
# moderate t-statistics and p-values
prolfqua::ContrastsModerated$undebug("get_contrasts")
contrasts <-  contrasts$merged |> prolfqua::ContrastsModerated$new() 
## visualize results in panels
pl <- contrasts$get_Plotter() 
panelA <- pl$histogram()$p.value + ggplot2::labs(tag = "A")                                                        
panelB <- pl$volcano()$FDR +
   ggplot2::theme(legend.position = "bottom") +
   ggplot2::labs(tag = "B")

gridExtra::grid.arrange(panelA, panelB, ncol = 2)
```

>>>>

The linear model below explain the observed protein abundances using the explanatory variables `condition_` and `batch_` and the interaction term  `condition_:batch_`:

```{r, echo = TRUE}
data_Yeast2Factor <- prolfqua::prolfqua_data("data_Yeast2Factor")
mlm <- prolfqua::strategy_lm(
  transformedIntensity ~ condition_ + batch_ + condition_:batch_
  )
mm <- prolfqua::build_model(data_Yeast2Factor$data, mlm, subject = "protein_Id")
```

The treatment difference among the groups `Ethanol` and `Glucose` defined by the factor `condition`, and similarly among the groups defined by `batch_`, can be specified.

@@

Furthermore, the treatment differences among `Ethanol` and `Glucose` within the batches `p2370` and `p2691` can be specified.

@@

Finally, we can examine if the treatment differences in both batches are the same or are different.

@@

The code to compute the array of weights $c$ used to multiply the model coefficient $\beta$, to obtain the effect size estimates, and the variance of the estimate, is shown below:

@@

Table \@ref(tab:contrcoefficient) shows weights $c$ for each of the five contrasts we specified.

@@

This example is taken from the `r BiocStyle::Githubpkg("fgcz/prolfqua")` package vignette file.
Given the weights $c$ contrasts from model parameter estimates obtained with `r BiocStyle::Biocpkg('proDA')`, or `r BiocStyle::Biocpkg('msqrob2')` can be determined, by this enabling us to call these packages using the same application programming interface (API).

@@

## compute contrasts from linear model
contr <- prolfqua::Contrasts$new(models, contrasts) 
# compute contrasts using imputation imputation
conI <- prolfqua::ContrastsSimpleImpute$new( lfqp, contrasts)
# merge contrasts, to obtain differences for all proteins
contrasts <- prolfqua::merge_contrasts_results(prefer = contr, add = conI)
# moderate t-statistics and p-values
prolfqua::ContrastsModerated$undebug("get_contrasts")
contrasts <-  contrasts$merged |> prolfqua::ContrastsModerated$new() 
## visualize results in panels
pl <- contrasts$get_Plotter() 
histpval <- pl$histogram()$p.value + ggplot2::labs(tag = "A")               maplot <- pl$ma_plot(rank = TRUE) + ggplot2::labs(tag = "C")
volcano <- pl$volcano()$FDR +
   ggplot2::theme(legend.position = "bottom") +
   ggplot2::labs(tag = "B")

gridExtra::grid.arrange(histpval, maplot, volcano, layout_matrix = rbind(c(1,3),c(2,3)))
```
